\documentclass{ximera}

\input{../preamble.tex}

\author{Carl Stitz \and Jeff Zeager \and Matthew Carr}
\license{CC-By-SA-NC}
\acknowledgement{http://www.stitz-zeager.com/}

\begin{document}
\begin{exercise}


\outcome{Find domain and range.}
\outcome{Define and work with inverse functions.}
\outcome{Determine where a function is positive or negative}
\outcome{Find the intervals where a function is increasing or decreasing}

\tag{domain}
\tag{inverse}
\tag{increasing}
\tag{decreasing}

Now use the graph of $y = f(x)$ given below to answer the questions.
\begin{image}
  \begin{tikzpicture}
    \begin{axis}[
        xmin=-5, xmax=5, ymin=-6,ymax=6,    
        unit vector ratio*=1 1 1,
        axis lines =middle, xlabel=$x$, ylabel=$y$,
        every axis y label/.style={at=(current axis.above origin),anchor=south},
        every axis x label/.style={at=(current axis.right of origin),anchor=west},
        xtick={-5,...,5}, ytick={-5,...,5},
        %xticklabels={-4,,-2,,0,,2,,4,,6}, yticklabels={,-2,,0,,2,,4,,6,,8,,10},
        grid=major,width=4in,
        grid style={dashed, gridColor},
      ]
      \addplot[color=penColor,very thick,smooth,domain=-4:4] {5*sin(x*180/4)};
      \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(2,5)};  %% open hole
      \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(2,3)};  %% closed hole
      
    \end{axis}
  \end{tikzpicture}
\end{image}
\begin{multicols}{2}
\begin{enumerate}
\item  Find the domain of $f$. \label{usesecondfuncgraphfirst}
\item  Find the range of $f$.
\item  Determine $f(2)$.
\item  Solve $f(x) = -5$.
\item  List the $x$-intercepts from left to right. If none exist, write DNE.
\item  List the $y$-intercepts from bottom to top. If none exist, write DNE.
\item  List the zeros of $f$ from left to right. If none exist, write DNE.
\item  Solve $f(x) \leq 0$ using intervals written from left to right.
\item  Find the number of solutions to $f(x) = 3$.
\item  List the intervals where $f$ is increasing written from left to right. If none exist, write DNE.
\item  List the intervals where $f$ is decreasing written from left to right. If none exist, write DNE.
\item  Restricting to the domain $[-2,2]$ estimate $f^{-1}(2)$.
\item  Restricting to the domain $[2,4]$ estimate $f^{-1}(2)$.
\end{enumerate}
\end{multicols}

Answer:

\begin{multicols}{2}
\begin{enumerate}
\item  $[\answer{-4},\answer{4}]$
\item  $[\answer{-5},\answer{5})$
\item  $f(2)=\answer{3}$
\item  $x=\answer{-2}$
\item  $\answer{-4}$, $\answer{0}$ and $\answer{4}$
\item  $\answer{0}$
\item  $\answer{-4}$, $\answer{-1}$ and $\answer{1}$
\item  $[\answer{-4},\answer{-1}]$ and $[\answer{1},\answer{3}]$
\item  There are $\answer{3}$ solutions.
\item  $[\answer{-2},\answer{2})$
\item  $[\answer{-3},\answer{0}]$ and $[\answer{2},\answer{3}]$
\item  $f^{-1}(2)\approx\answer{\frac{1}{2}}$
\item  $f^{-1}(3)\approx\answer{\frac{7}{2}}$
\end{enumerate}
\end{multicols}

\end{exercise}
\end{document}