\documentclass{ximera}

\input{preamble.tex}

\author{Gregory Hartman \and Matthew Carr}
\license{Creative Commons 3.0 By-NC}
\acknowledgement{https://github.com/APEXCalculus}

\begin{document}
\begin{exercise}

\outcome{Interpret an optimization problem as the procedure used to make a system or design as effective or functional as possible.}
\outcome{Set up an optimization problem by identifying the objective function and appropriate constraints.}
\outcome{Solve optimization problems by finding the appropriate absolute extremum.}
\outcome{Solve basic word problems involving maxima or minima.}

\tag{extrema}
\tag{constrained optimization}
\tag{first derivative test}
\tag{second derivative test}

Find the absolute maximum product of two numbers (not necessarily integers) whose sum is $100$. Write DNE is no absolute maximum exists. \[\answer{2500}\]


\begin{hint}
We want to find the absolute maximum the product of two numbers (call them $x$ and $y$). Thus, we want to maximize $xy$ subject to the constraint that $x+y=100$.
\end{hint}
\begin{hint}
Solving for $y$, we see that $y=100-x$. Then $xy=-x^2+100x$. We must find the absolute maximum of this product. 
\end{hint}
\begin{hint}
If there is an absolute maximum, then, since $-x^2+100x$ is defined everywhere, the absolute maximum must occur where there is a local maximum. Hence, we find the critical points of $-x^2+100x$: $\frac{d}{dx}(-x^2+100x)=0$ and $\frac{d}{dx}(-x^2+100x)=100-2x$, hence, solving for $x$ in $100-2x=0$, we have $x=50$. So $x=50$ and $y=50$.

But is this a maximum or a minimum? By the second derivative test, $\frac{d^2}{dx^2}(-x^2+100x)=-2$ and $-2<0$, so this is a maximum. We see by inspection that it is an absolute maximum because the function does not attain a larger value.

Thus, $xy$ is maximized with $xy=2500$. 
\end{hint}





\end{exercise}
\end{document}