\documentclass{ximera}

\input{../preamble.tex}

\author{Gregory Hartman \and Matthew Carr}
\license{Creative Commons 3.0 By-NC}
\acknowledgement{https://github.com/APEXCalculus}

\begin{document}
\begin{exercise}

\outcome{Interpret an optimization problem as the procedure used to make a system or design as effective or functional as possible.}
\outcome{Set up an optimization problem by identifying the objective function and appropriate constraints.}
\outcome{Solve optimization problems by finding the appropriate absolute extremum.}
\outcome{Solve basic word problems involving maxima or minima.}

\tag{extrema}
\tag{constrained optimization}
\tag{first derivative test}
\tag{second derivative test}

Find the absolute maximum of a sum of two numbers, each of which is in $[0,300]$ whose product is $500$.\[\answer{\frac{905}{3}}\]

\begin{hint}
If the two numbers are $x$ and $y$, then $xy=500$, so $y=\frac{500}{x}$, so we are left with the job of finding the absolute maximum of $x+\frac{500}{x}$ subject to the constraint that $x$ lie in $[0,300]$ (since $\frac{500}{x}\le300$ for all $x$ in $[0,300]$, we needn't worry that $y$ lies outside of $[0,300]$).
\end{hint}
\begin{hint}
The only critical point of $x+\frac{500}{x}$ that lies in $[0,300]$ is a minimum. Since $[0,300]$ is a closed interval (i.e., it contains its endpoints), what can you say about where the maximum of $x+\frac{500}{x}$ must occur?
\end{hint}
\begin{hint}
Recall that extrema can occur at critical points or at the endpoints of a closed interval, such as $[0,300]$. 
\end{hint}
\begin{hint}
If $xy=500$, then $x$ cannot be $0$, so check the other endpoint.
\end{hint}
\begin{hint}
If $x=300$, then $300+\frac{500}{300}=\frac{905}{3}$ must be the maximum. Indeed, it is the absolute maximum, for if $x<300$, then clearly $x+\frac{500}{x}<\frac{905}{3}$.
\end{hint}
\end{exercise}
\end{document}
