\documentclass{ximera}
\input{../preamble.tex}
\author{Steven Gubkin}
\license{Creative Commons 3.0 By-NC}
\begin{document}
\begin{exercise}

\outcome{Define accumulation functions.}
\outcome{Calculate and evaluate accumulation functions.}
\outcome{State the First Fundamental Theorem of Calculus.}
\outcome{Take derivatives of accumulation functions using the First Fundamental Theorem of Calculus.}
\outcome{Use accumulation functions to find information about the original function.}
\outcome{Understand the relationship between the function and the derivative of its accumulation function.}

\tag{integral}

Usually we use the second fundamental theorem of calculus to compute
definite integrals.  In other words, we turn a hard geometric problem
(finding the area under a curve) into a relatively easy algebra
problem (evaluating an antiderivative, and evaluating at the
endpoints).

In this problem we will turn a hard antiderivative problem into a relatively easy geometry problem using the first fundamental theorem of calculus.

Using the picture below, find an algebraic expression for $\int_{-1}^x
\sqrt{1-t^2} \d t$.  Verify that this is an antiderivative of
$\sqrt{1-x^2}$ by taking the derivative.


%BADBAD add a filled arc from (-1,0) to (0.5,sqrt(3)/2).  Label the two regions as A and B.

\begin{image}
  \begin{tikzpicture}
    \begin{axis}[
        width=5in,
        height=3in,
        xmin=-1.2, xmax=1.2,ymin=-0.2,ymax=1.2,domain=-2:2,
        axis lines =center, xlabel=$x$, ylabel=$y$,
        every axis y label/.style={at=(current axis.above origin),anchor=south},
        every axis x label/.style={at=(current axis.right of origin),anchor=west},
        axis on top,
    ] 
      \addplot [draw=none, fill=penColor2!50!white,domain=0:0.5] {1.732*x} \closedcycle;
      \addplot [penColor,very thick,domain=0:pi] ({cos(x*360/2/pi)},{sin(x*360/2/pi)});
  \end{axis}
  \end{tikzpicture}
\end{image}

\begin{prompt}
The area of region $A$, as a function of $x$, is $\answer{\pi - \arccos(x)}$

The area of region $B$, as a function of $x$, is $\answer{\frac{1}{2}x \sqrt{1-x^2}}$

So $\int_{-1}^x \sqrt{1-t^2} \d t = \answer{\pi - \arccos(x)+\frac{1}{2}x \sqrt{1-x^2}}$
\end{prompt}

\end{exercise}
\end{document}
